---
title: "R_Project2"
author: "Kai Yeung"
date: "2025-06-05"
output: html_document
---

```{r}
library(ggplot2)
library(zoo)

vaccine_admin <- read.csv("covid-19-progress-county-051021.csv")

summary(vaccine_admin)
```

?trimws
```{r}
vaccine_admin$county <- trimws(vaccine_admin$county, which = "both")

vaccine_admin$administered_date <- as.Date(vaccine_admin$administered_date, format = "%Y-%m-%d")
```

## Calculating the % of vaccinated
## using subset()

```{r}
vaccine_CA <- subset(vaccine_admin, county == "All CA Counties")
```

## CA population

```{r}
CA_pop <- c(39466917)

##calculating % of people vaccinated with a least one dose
vaccine_CA$partial_percent <- (vaccine_CA$cumulative_at_least_one_dose/CA_pop)*100

#rounding the % to 2 digits.
vaccine_CA$partial_percent <- round(vaccine_CA$partial_percent, digits = 2)

##For fully vaccinated and combining the two steps together above
vaccine_CA$fully_percent <- round(((vaccine_CA$cumulative_fully_vaccinated/CA_pop)*100), digits = 2)
```

Question: Take a look at the new percentages columns in the dataset (particularly the last one). Compare it to the percentages in the pdf (49.1% of residents have received at least one dose and 35.7% are fully vaccinated). Out percentages are close, but why do you think the calculations might differ from each other?

Answer: We are using the subset we created?

