---
title: "R_Project2"
author: "Kai Yeung"
date: "2025-06-05"
output: html_document
---

```{r}
library(ggplot2)
library(zoo)

vaccine_admin <- read.csv("covid-19-progress-county-051021.csv")

summary(vaccine_admin)
```

?trimws
```{r}
vaccine_admin$county <- trimws(vaccine_admin$county, which = "both")

vaccine_admin$administered_date <- as.Date(vaccine_admin$administered_date, format = "%Y-%m-%d")
```

## Calculating the % of vaccinated
## using subset()

```{r}
vaccine_CA <- subset(vaccine_admin, county == "All CA Counties")
```

## CA population

```{r}
CA_pop <- c(39466917)

##calculating % of people vaccinated with a least one dose
vaccine_CA$partial_percent <- (vaccine_CA$cumulative_at_least_one_dose/CA_pop)*100

#rounding the % to 2 digits.
vaccine_CA$partial_percent <- round(vaccine_CA$partial_percent, digits = 2)

##For fully vaccinated and combining the two steps together above
vaccine_CA$fully_percent <- round(((vaccine_CA$cumulative_fully_vaccinated/CA_pop)*100), digits = 2)
```

Question: Take a look at the new percentages columns in the dataset (particularly the last one). Compare it to the percentages in the pdf (49.1% of residents have received at least one dose and 35.7% are fully vaccinated). Out percentages are close, but why do you think the calculations might differ from each other?

Answer: different population data

```{r}
## show the column names in the dataset
names(vaccine_CA)

# Plotting the initial graph with the first layer for the partially vaccinated percentages
ggplot(data = vaccine_CA, aes(x = administered_date, y = partial_percent)) +
  geom_area(fill = "darkolivegreen3")

# Adding the second layer with the fully vaccinated percentages
ggplot(data = vaccine_CA, aes(x = administered_date, y = partial_percent)) +
  geom_area(fill = "darkolivegreen3") + 
  geom_area(aes(x = administered_date, y = fully_percent), fill="forestgreen")

# Here is an example of changing the opacity of the graph with the aesthetic parameter alpha
ggplot(data = vaccine_CA, aes(x = administered_date, y = partial_percent)) +
  geom_area(fill = "darkolivegreen3", alpha = 0.4) + 
  geom_area(aes(x = administered_date, y = fully_percent), fill="forestgreen", alpha = 0.4)
```


```{r horizontal line}

# Try changing the line type with something different

ggplot(data = vaccine_CA, aes(x = administered_date, y = partial_percent)) +
  geom_area(fill="darkolivegreen3", alpha = 0.4) + 
  geom_area(aes(x= administered_date, y = fully_percent), fill="forestgreen", alpha = 0.4) +
  ylim(0, 100) +
  geom_hline(yintercept = 80, linetype='dotdash', col = "olivedrab")

```

```{r summary overlay}

# Use the function rollmean() to calculate rolling averages over 7 days
vaccine_CA$wk_avg <- zoo::rollmean(vaccine_CA$total_doses, k = 7, fill = NA)

# Round the number to the nearest whole number
vaccine_CA$wk_avg <- round(vaccine_CA$wk_avg, digits = 0)

# Because the data has NAs, a warning message will appear: "Removed 6 row(s) containing missing values (geom_path)"
# So you can na.rm = T to remove the NAs from the plot
ggplot(data = vaccine_CA, aes(x = administered_date, y = total_doses)) +
  geom_col(fill = "darkolivegreen3", alpha = 0.4) +
  scale_y_continuous() +
  geom_line(data = vaccine_CA, aes(x = administered_date, y = wk_avg), na.rm = T, color = "olivedrab")

```

```{r subset county}
# Fill in the county of your chose to subset
vaccine_ala <- subset(vaccine_admin, county == "Alameda")

# Fill in the county's population
ala_pop <- c(1682353)

# Calculate the percentage the percentage of those who have had at least one dose in that county

vaccine_ala$partial_percent <- round(((vaccine_ala$cumulative_at_least_one_dose/ala_pop)*100), digits = 2)

# Calculate the percentage of those who are fully vaccinated in that county
vaccine_ala$fully_percent <- round(((vaccine_ala$cumulative_fully_vaccinated/ala_pop)*100), digits = 2)

```

#plot 1: % for Alameda County

```{r horizontal line}

# alameda

alameda_Pecentage <- ggplot(data = vaccine_ala, aes(x = administered_date, y = partial_percent)) +
                        geom_area(fill="darkolivegreen3", alpha = 0.4) + 
                        geom_area(aes(x= administered_date, y = fully_percent), fill="forestgreen", alpha = 0.4) +
                        ylim(0, 100) +
                        geom_hline(yintercept = 80, linetype='dotdash', col = "olivedrab")+
                        ggtitle("Percentage of Alameda Residents Who are Partially or Fully Vaccinated")+
                        xlab("Administered Date") + ylab("Percentage")

print(alameda_Pecentage)
pdf("Alameda Vaccinated Percentage.pdf")

print(alameda_Pecentage)
```

```{r}



# Plot 2: Doses administered by Alameda county 


# Use the function rollmean() to calculate rolling averages over 7 days
vaccine_ala$wk_avg <- zoo::rollmean(vaccine_ala$total_doses, k = 7, fill = NA)

# Round the number to the nearest whole number
vaccine_ala$wk_avg <- round(vaccine_ala$wk_avg, digits = 0)

# Because the data has NAs, a warning message will appear: "Removed 6 row(s) containing missing values (geom_path)"
# So you can na.rm = T to remove the NAs from the plot
Alameda_Dose <- ggplot(data = vaccine_ala, aes(x = administered_date, y = total_doses)) +
  geom_col(fill = "darkolivegreen3", alpha = 0.4) +
  scale_y_continuous() +
  geom_line(data = vaccine_ala, aes(x = administered_date, y = wk_avg), na.rm = T, color = "olivedrab")+
  ggtitle("Doses administered by Alameda County")+
  ylab("Total Doses") + xlab("Administered Date")+
  theme(legend.position = "right")

print(Alameda_Dose)
```

